\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-07-25 redaction/dhp_july2017.tex
modifsJH250717
Début révision texte et OA.

\documentclass[a4paper,11pt]{article}
%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
%%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
%J'ai essayé de débuger : je n'y suis pas arrivé
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage[tableposition=top]{caption}
\usepackage{ bbold }
\usepackage{ upgreek }
\usepackage{comment}
\includecomment{commentGD}
\usepackage[draft]{todonotes}
%\usepackage[disable]{todonotes}
%%GDPour cacher les notes, \usepackage[disable]{todonotes}

@@

\begin{abstract}
Following \cite{samuelson1954}, standard models of international trade have usually modelled trade costs as an ad-valorem tax equivalent. However, many common empirical facts point out the importance of additive costs. This paper provides new empirical results that support this view. To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and ocean). Three main results emerge. First, we show the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime transport respectively. Second, taking additive costs into account significantly improves the fit of the modelling of transport costs. Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs. Allowing for a share of additive costs varying in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs rather than to trade pattern composition effects. In all three aspects, our results point the importance of the additive component in accounting for international transport costs.

@@

\begin{table}[htbp]
  \centering
  \caption{Air: Measures of Goodness-of-fit (3 digits)}
  \footnotesize{
\begin{center}
    \begin{tabular}{l|cccccc|c}
    \hline \hline
    Year  & 1974  & 1980  & 1990  & 2000  & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
    \multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
    Term I only & 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
    Terms A \& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
    \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
    Term I only & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & 0.85 \\
    Terms A \& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
    Term I only & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{88191.87} & 70498.1 \\
    Terms A \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{80873.72} & 62285.0 \\ \hline
    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
    Term I only & -17530.5 & -20253.5 & -29977.8 & -43341.3 & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-43692.9} & -34888.6 \\
    Terms A \& I & -15125.6 & -17263.2 & -25393.5 & -36788.4 & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-39751.9} & -30508.3 \\
    LL ratio & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
    nb of restrictions & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{427} & 402 \\
    p-value & 0.00 & 0.00 & 0.00 & 0.00 & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00 \\
    \hline \hline
 \end{tabular}%
    \end{center}}
  \label{tab:good_fit_air}%
 \parbox[l]{12cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion. $R^{2}$ between the log of predicted ratio and the log of the observed ratio. For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products. The mean statistics calculated as the average value over all years. }}
\end{table}%

\begin{table}[htbp]
  \centering
  \caption{Vessel: Measures of Goodness-of-fit (3 digits)}
    \footnotesize{
\begin{center}
\begin{tabular}{l|cccccc|c}
\hline \hline
Year  & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & 2010  & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
\multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
Term I only & \multicolumn{1}{c}{0.450} & \multicolumn{1}{c}{0.415} & \multicolumn{1}{c}{0.456} & \multicolumn{1}{c}{0.401} & 0.350 & \multicolumn{1}{c}{0.339} & 0.39 \\
Terms A \& I & \multicolumn{1}{c}{0.612} & \multicolumn{1}{c}{0.575} & \multicolumn{1}{c}{0.590} & \multicolumn{1}{c}{0.571} & 0.491 & \multicolumn{1}{c}{0.462} & 0.56 \\ \hline
\multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
    Term I only & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & 0.74  & \multicolumn{1}{c}{0.76} & 0.66 \\
    Terms A \& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
    Term I only & \multicolumn{1}{c}{33328.8} & \multicolumn{1}{c}{33010.3} & \multicolumn{1}{c}{51142.6} & \multicolumn{1}{c}{71365.9} & 84789.9 & \multicolumn{1}{c}{88191.9} & 57848.6 \\
    Terms A \& I & \multicolumn{1}{c}{27331.5} & \multicolumn{1}{c}{28067.3} & \multicolumn{1}{c}{43664.7} & \multicolumn{1}{c}{60475.9} & 76161.3 & \multicolumn{1}{c}{80873.7} & 49682.3 \\ \hline
    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
    Term I only & \multicolumn{1}{c}{-16287.4} & \multicolumn{1}{c}{-16129.1} & \multicolumn{1}{c}{-25169.3} & \multicolumn{1}{c}{-35263.9} & -41998.9 & \multicolumn{1}{c}{-43692.9} & -28534.3 \\
    Terms A \& I & \multicolumn{1}{c}{-12985.8} & \multicolumn{1}{c}{-13353.7} & \multicolumn{1}{c}{-21171.4} & \multicolumn{1}{c}{-29491.0} & -37418.7 & \multicolumn{1}{c}{-39751.9} & -24151.3 \\
    LL ratio & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & 9160.56 & \multicolumn{1}{c}{7882.15} & 8766.0 \\
    nb of restrictions & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & 424   & \multicolumn{1}{c}{427} & 417 \\
    p-value& \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00  & \multicolumn{1}{c}{0.00} & 0.00 \\
    \hline \hline
    \end{tabular}%
    \end{center}}
  \label{tab:good_fit_vessel}%
  \parbox[l]{12cm}{\tiny{Notes: SER = Standard Error of regression; AIC = Akaike Information Criterion. $R^{2}$ between the log of predicted ratio and the log of the observed ratio. For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products. The mean statistics calculated as the average value over all years. }}
\end{table}%

>>>>

\documentclass[a4paper,11pt]{article}
%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
%%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
%J'ai essayé de débuger : je n'y suis pas arrivé
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage[tableposition=top]{caption}
\usepackage{amsfonts}
%\usepackage{bbold}
% JH : impossible de compiler avec le package bbold, remplacé par amsfonts
\usepackage{upgreek}
\usepackage{comment}
\includecomment{commentGD}
\usepackage[draft]{todonotes}
%\usepackage[disable]{todonotes}
%%GDPour cacher les notes, \usepackage[disable]{todonotes}

@@

\begin{abstract}
Following \cite{samuelson1954}, standard models of international trade have usually modelled trade costs as an ad-valorem tax equivalent. However, many common empirical facts point out the importance of additive costs. This paper provides new empirical results that support this view. To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel). Three main results emerge. First, we show the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and maritime vessel transport respectively. Second, taking additive costs into account significantly improves the fit of the modelling of transport costs. Third, we show the importance of integrating the additive component in accounting for the time trend of international transport costs. Allowing for a share of additive costs varying in product/country/time dimension, we obtain that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{ceteris paribus} transport costs rather than to trade pattern composition effects. In all three aspects, our results point the importance of the additive component in accounting for international transport costs.

@@

\begin{table}[htbp]
  \centering
  \caption{Air: Measures of Goodness-of-fit (3 digits)}
  \footnotesize{
\begin{center}
    \begin{tabular}{l|cccccc|c}
    \hline \hline
    Year  & 1974  & 1980  & 1990  & 2000  & \multicolumn{1}{c}{2010} & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
    \multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
    Model (A)& 0.30  & 0.27  & 0.25  & 0.32  & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.34} & 0.31 \\
    Model (B)\& I & 0.59  & 0.65  & 0.63  & 0.64  & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} & 0.60 \\ \hline
    \multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
     Model (A) & 0.79  & 0.86  & 0.81  & 0.84  & \multicolumn{1}{c}{0.86} & \multicolumn{1}{c}{0.92} & 0.85 \\
    Model (B)\& I & 0.67  & 0.71  & 0.67  & 0.70  & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.85} & 0.73 \\ \hline
   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
    Model (A) & 35674.98 & 41170.98 & 60715.58 & 87492.55 & \multicolumn{1}{c}{102297.66} & \multicolumn{1}{c}{88191.87} & 70498.1 \\
    Model (B) \& I & 31387.29 & 35738.39 & 52098.91 & 74954.88 & \multicolumn{1}{c}{95887.05} & \multicolumn{1}{c}{80873.72} & 62285.0 \\ \hline
    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
    Model (A) & -17530.5 & -20253.5 & -29977.8 & -43341.3 & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-43692.9} & -34888.6 \\
    Model (B) & -15125.6 & -17263.2 & -25393.5 & -36788.4 & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-39751.9} & -30508.3 \\
    LL ratio & 4809.7 & 5980.6 & 9168.7 & 13105.7 & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{7882.1} & 8760.69 \\
    nb of restrictions & 355   & 369   & 393   & 426   & \multicolumn{1}{c}{426} & \multicolumn{1}{c}{427} & 402 \\
    p-value & 0.00 & 0.00 & 0.00 & 0.00 & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00 \\
    \hline \hline
 \end{tabular}%
    \end{center}}
  \label{tab:good_fit_air}%
 \parbox[l]{12cm}{\tiny{Notes: Model (A) = with only ad-valorem transport costs. Model (B) = with additive \& Ad-Valorem transport costs. SER = Standard Error of regression; AIC = Akaike Information Criterion. $R^{2}$ between the log of predicted ratio and the log of the observed ratio. For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products. The mean statistics calculated as the average value over all years. }}
\end{table}%

\begin{table}[htbp]
  \centering
  \caption{Vessel: Measures of Goodness-of-fit (3 digits)}
    \footnotesize{
\begin{center}
\begin{tabular}{l|cccccc|c}
\hline \hline
Year  & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1980} & \multicolumn{1}{c}{1990} & \multicolumn{1}{c}{2000} & 2010  & \multicolumn{1}{c}{2013} & Mean stat \\ \hline
\multicolumn{8}{l}{\bf{$R^2$} }\\ \hline
Model (A)& \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.42} & \multicolumn{1}{c}{0.46} & \multicolumn{1}{c}{0.40} & 0.35 & \multicolumn{1}{c}{0.34} & 0.39 \\
Terms A \& I & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.57} & 0.49 & \multicolumn{1}{c}{0.46} & 0.56 \\ \hline
\multicolumn{8}{l}{\textbf{SER}  }  \\ \hline
    Model (A) & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.59} & \multicolumn{1}{c}{0.65} & 0.74  & \multicolumn{1}{c}{0.76} & 0.66 \\
    Model (B)\& I & \multicolumn{1}{c}{0.48} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.55} & 0.66  & \multicolumn{1}{c}{0.68} & 0.57 \\ \hline
   \multicolumn{8}{l}{\textbf{AIC criteria}}  \\ \hline
   Model (A) & \multicolumn{1}{c}{33328.8} & \multicolumn{1}{c}{33010.3} & \multicolumn{1}{c}{51142.6} & \multicolumn{1}{c}{71365.9} & 84789.9 & \multicolumn{1}{c}{88191.9} & 57848.6 \\
   Model (B) & \multicolumn{1}{c}{27331.5} & \multicolumn{1}{c}{28067.3} & \multicolumn{1}{c}{43664.7} & \multicolumn{1}{c}{60475.9} & 76161.3 & \multicolumn{1}{c}{80873.7} & 49682.3 \\ \hline
    \multicolumn{8}{l}{\textbf{Log-likelihood}} \\ \hline
    Model (A) & \multicolumn{1}{c}{-16287.4} & \multicolumn{1}{c}{-16129.1} & \multicolumn{1}{c}{-25169.3} & \multicolumn{1}{c}{-35263.9} & -41998.9 & \multicolumn{1}{c}{-43692.9} & -28534.3 \\
    Model (B) & \multicolumn{1}{c}{-12985.8} & \multicolumn{1}{c}{-13353.7} & \multicolumn{1}{c}{-21171.4} & \multicolumn{1}{c}{-29491.0} & -37418.7 & \multicolumn{1}{c}{-39751.9} & -24151.3 \\
    LL ratio & \multicolumn{1}{c}{6603.28} & \multicolumn{1}{c}{5550.96} & \multicolumn{1}{c}{7995.88} & \multicolumn{1}{c}{11545.98} & 9160.56 & \multicolumn{1}{c}{7882.15} & 8766.0 \\
    nb of restrictions & \multicolumn{1}{c}{393} & \multicolumn{1}{c}{395} & \multicolumn{1}{c}{411} & \multicolumn{1}{c}{436} & 424   & \multicolumn{1}{c}{427} & 417 \\
    p-value& \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & 0.00  & \multicolumn{1}{c}{0.00} & 0.00 \\
    \hline \hline
    \end{tabular}%
    \end{center}}
  \label{tab:good_fit_vessel}%
  \parbox[l]{12cm}{\tiny{Notes: Model (A) = with only ad-valorem transport costs. Model (B) = with additive \& Ad-Valorem transport costs. SER = Standard Error of regression; AIC = Akaike Information Criterion. $R^{2}$ between the log of predicted ratio and the log of the observed ratio. For the LL ratio test, the number of restrictions is equal to the number of parameters estimated, i.e., the number of partner countries plus the number of products. The mean statistics calculated as the average value over all years. }}
\end{table}%
